\begin{problem}{CSPs: Midterm 1 Staff Assignments}

The CS188 Midterm is coming up, and the CS188 staff has yet to write the test. There are a total of 6 questions on the exam and each question will cover a topic. Here is the format of the exam:

\begin{itemize}
\item Q1. Search
\item Q2. Games
\item Q3. CSPs
\item Q4. True/False
\item Q5. Agents
\item Q6. Short Answer
\end{itemize}

There are 3 people on the course staff who will be writing questions: John, Kevin, and Anna. Each question will have only one TA assigned to writing it. However, the staff are pretty quirky and want the following constraints to be satisfied:
\begin{itemize}
\item Kevin will only work on Search, Games and CSPs.
\item Anna is very odd, so she can only contribute to an odd-numbered question.
\item The same TA must write both the True/False and the Short Answer question.
\item No TA may work on two consecutive questions.
\end{itemize}
\vspace{0.5cm}

\newpage

\begin{question}
Formulate this problem as a CSP problem in which there is one variable per class, stating the domains, and constraints. Constraints should be specified formally and precisely, but may be implicit rather than explicit.

\begin{subquestion}[7] List the variables and domains after unary constraints are applied.\\\\
\fbox{\begin{minipage}[t][6cm][t]{13cm}
\AnswerOneAi 
\end{minipage}}
\end{subquestion}

\begin{subquestion}[8] List the constraints.\\\\
\fbox{\begin{minipage}[t][6cm][t]{13cm}
\AnswerOneAii
\end{minipage}}
\end{subquestion}
\end{question}

\begin{question}[10] Draw the constraint graph associated with your CSP.\\\\
\fbox{\begin{minipage}[t][7cm][t]{13cm}
\AnswerOneB
\end{minipage}}
\end{question}

\end{problem}