\documentclass{article}
\usepackage{listings}
\title{Task 4 Design}

\author{Alexander Wang}

\begin{document}
\lstset{language=Python}
\maketitle

\section*{\underline{Task 4: Communicator}}

\subsection*{Overview}
In order to ensure that only one speaker and listener are transferring at a time, we will use a lock and a condition variable. When \texttt{listen()} gets called, it will acquire the lock, increase the number of waiting listeners, and check to see if the number of waiting speakers is greater than one. If so, \texttt{listen()} will \texttt{signal()} and then \texttt{wait()} for the speaker to transfer the word; otherwise, it will simply call \texttt{wait()}. Upon the waking up from the second \texttt{wait()}, \texttt{listen()} will release the lock and return the word. When \texttt{speak()} gets called, it will acquire the lock, increase the number of waiting speakers, and check to see if the number of waiting listeners is greater than one. If so, \texttt{speak()} will decrease the number of waiting speakers, \texttt{signal()} a listener, and release the lock.

\subsection*{Correctness Constraints}
\begin{itemize}
\item A speaker and listener are never both waiting
\item Exactly one listener receives a word given by any speaker
\end{itemize}

\subsection*{Declarations}
\texttt{Lock lock} - Ensure only one communicator is modifying condition variables at any given time. \\
\texttt{int speakersWaiting} - Allow listener to determine if a speaker is available. \\
\texttt{int listenersWaiting} - Allow speaker to determine if a listener is available. \\
\texttt{int currentWord} - 32-bit word that is to be replaced by the speaker in transfer. \\
\texttt{Condition okToTransfer} - Conditional variable to let speaker send. \\
\texttt{Condition okToReceive} - Conditional variable to let receiver receive. \\
\texttt{Condition transferMade} - Conditional variable to transfer.

\subsection*{Descriptions}

\begin{lstlisting}[frame=single]
Lock lock
word = nil
speakersWaiting = 0
listenersWaiting = 0

listen():
    acquire lock
    while speakersWaiting is 0:
        increment listenersWaiting
        wait on okToReceive and release lock
        decrement listenersWaiting
    signal okToTransfer
    wait on transferMade and release lock again
    return word

speak(myWord):
    acquire lock
    while listenersWaiting is 0:
        increment speakersWaiting
        wait on okToTransfer and release lock
        decrement speakersWaiting
    word = myWord
    signal transferMade
    release lock
\end{lstlisting}

\subsection*{Testing Plan}
We plan to test Communicator by including test methods.
\begin{itemize}
\item Initialize 4 speakers each with unique words
\subitem Call listen one by one and see if the 4 words are returned as expected
\subitem Check that both speakersWaiting and listenersWaiting are 0
\item Initialize 4 listeners
\subitem Call speak one by one with unique words and see that each is returned as expected
\subitem Check that both speakersWaiting and listenersWaiting are 0
\item Choose either a speaker or listener with 50 percent probability and keep track of how many of each there are.
\subitem At the end, get the absolute value of speakers minus listeners created and ensure that it equals speakersWaiting + listenersWaiting
\subitem Get the minimum of the two and ensure that it equals the number of words returned
\end{itemize}

\end{document}