\q{20}{Decision Trees}

You are given the following data sets

\begin{tabular}{c c c c c c}
(i)
&
\begin{tabular}{|c | c | c |}
\hline
$X_1$ & $X_2$ & $Y$\\
\hline
1 & 1 & +\\
\hline
4 & 5 & +\\
\hline
4 & 5 & -\\
\hline
5 & 5 & +\\
\hline
\end{tabular}
&
(ii)
&
\begin{tabular}{|c | c | c |}
\hline
$X_1$ & $X_2$ & $Y$\\
\hline
1 & 1 & +\\
\hline
4 & 3 & +\\
\hline
4 & 5 & -\\
\hline
5 & 5 & +\\
\hline
\end{tabular}
&
(iii)
&
\begin{tabular}{|c | c | c |}
\hline
$X_1$ & $X_2$ & $Y$\\
\hline
1 & 1 & +\\
\hline
4 & 2 & -\\
\hline
4 & 5 & -\\
\hline
5 & 5 & +\\
\hline
\end{tabular}
\end{tabular}


\begin{question}{}
  [3 pts] Which data sets are linearly separable?
\end{question}
\begin{minipage}{\textwidth}
    \solution{\ \vspace{2cm}} {
    Data set (ii)
    }
\end{minipage}

\begin{question}{}
  [3 pts] Which data sets have label noise?
\end{question}
\begin{minipage}{\textwidth}
    \solution{\ \vspace{2cm}} {
    Data set (i)
    }
\end{minipage}

\begin{question}{}
  [3 pts] Which data sets can be fit exactly by a decision tree?
\end{question}
\begin{minipage}{\textwidth}
    \solution{\ \vspace{2cm}} {
    Data sets (ii) and (iii)
    }
\end{minipage}

\begin{question}{}
  [5 pts] A 1-decision-list is a decision tree in which the "yes" branch of every binary test is a leaf node. For a continuous attribute $X_j$, a test can be either $X_j > c$ or $X_j < c$. Continuous attributes can appear in multiple tests. Pick a data set and show a decision list that fits it exactly.
\end{question}
\begin{minipage}{\textwidth}
    \solution{} { 
    Data set (iii) in pseudocode: \begin{verbatim}
         if x1 < 4 : return +
    else if x2 < 5 : return -
    else if x1 < 5 : return -
    else           : return +
    \end{verbatim}
    }
\end{minipage}

\begin{question}{}
  [6 pts] In the absense of label noise, can any two-class data set in two dimensions be fit exactly by a decision list? Briefly explain why, or give a counterexample.
\end{question}
\begin{minipage}{\textwidth}
    \solution{} {
    \paragraph{} The data set $\{((-1,-1),+),((-1,1),-),((1,1),+),((1,-1),-)\}$ is a counterexample. No horizontal or vertical line (which are the only tests we can impose) creates a region of exactly one class.
    }
\end{minipage}


